Fitness effects of mutations to SARS-CoV-2 proteins

### 2023-08-07 paper/paper.tex
fixed editor comment 1 (typo)

\section{Methods}
\subsection{Code and data availability}
See the GitHub repository at \url{https://github.com/jbloomlab/SARS2-mut-fitness} for the computer code and processed data (eg, fitness estimates and mutation counts).
That repository contains a README with links to specific data files as well as a description of the computational pipeline.
See \url{https://github.com/jbloomlab/SARS2-mut-fitness/blob/main/results/aa_fitness/aa_fitness.csv} final estimates of amino-acid fitnesses across all clades; other intermediate data files are also provided in the GitHub repository.
The specific version of the repository used for this paper is tagged as ``bioRxiv-v2'' on GitHub (\url{https://github.com/jbloomlab/SARS2-mut-fitness/tree/bioRxiv-v2})
The pipeline is fully reproducible, and is run using \texttt{snakemake}~\citep{molder2021sustainable} with interactive plots rendered using \texttt{altair}~\citep{vanderplas2018altair}.

@@

We then performed quality control by ignoring any branch that met any of the following criteria:
\begin{itemize}
\item it had more than four nucleotide mutations;
\item it contained more than one nucleotide mutation that was a reversion to the Wuhan-Hu-1 reference sequence;
\item it contained more than one nucleotide mutation that was a reversion to the founder sequence for that clade as provided at \url{https://raw.githubusercontent.com/neherlab/SC2_variant_rates/7e738194a8c6592082f1caa9a6ca70cb68289790/data/clade_gts.json} by \cite{neher2022contributions};
\item it contained more than one nucleotide mutation to the same codon.
\end{itemize}
The rationale for the first exclusion is that highly mutated branches are often indicative of sequencing errors of viral evolution in chronically infected humans, neither of which correspond to the pattern of typical SARS-CoV-2 transmission in acute infections.
Because the virus's evolution is very densely sampled, only a small fraction of branches have more than four mutations (Figure~\ref{fig:nmut_counts}).
The rationale for the second and third exclusions is that excess reversions can arise from base-calling pipelines that erroneously call low-coverage sites as reference.
We ignore branches with multiple nucleotide mutations to the same codon (this is very rare) because as detailed below our method is only designed to make estimates for mutations that represent single-nucleotide changes from the clade founder.
Note also that the mutation-annotated tree does not include insertion or deletion mutations, and so we only consider (and make estimates for) point mutations.

>>>>

\section{Methods}
\subsection{Code and data availability}
See the GitHub repository at \url{https://github.com/jbloomlab/SARS2-mut-fitness} for the computer code and processed data (eg, fitness estimates and mutation counts).
That repository contains a README with links to specific data files as well as a description of the computational pipeline.
See \url{https://github.com/jbloomlab/SARS2-mut-fitness/blob/main/results/aa_fitness/aa_fitness.csv} for final estimates of amino-acid fitnesses across all clades; other intermediate data files are also provided in the GitHub repository.
The specific version of the repository used for this paper is tagged as ``bioRxiv-v2'' on GitHub (\url{https://github.com/jbloomlab/SARS2-mut-fitness/tree/bioRxiv-v2}).
The pipeline is fully reproducible, and is run using \texttt{snakemake}~\citep{molder2021sustainable} with interactive plots rendered using \texttt{altair}~\citep{vanderplas2018altair}.

@@

We then performed quality control by ignoring any branch that met any of the following criteria:
\begin{itemize}
\item it had more than four nucleotide mutations;
\item it contained more than one nucleotide mutation that was a reversion to the Wuhan-Hu-1 reference sequence;
\item it contained more than one nucleotide mutation that was a reversion to the founder sequence for that clade as provided at \url{https://raw.githubusercontent.com/neherlab/SC2_variant_rates/7e738194a8c6592082f1caa9a6ca70cb68289790/data/clade_gts.json} by \cite{neher2022contributions};
\item it contained more than one nucleotide mutation to the same codon.
\end{itemize}
The rationale for the first exclusion is that highly mutated branches are often indicative of sequencing errors or viral evolution in chronically infected humans, neither of which correspond to the pattern of typical SARS-CoV-2 transmission in acute infections.
Because the virus's evolution is very densely sampled, only a small fraction of branches have more than four mutations (Figure~\ref{fig:nmut_counts}).
The rationale for the second and third exclusions is that excess reversions can arise from base-calling pipelines that erroneously call low-coverage sites as reference.
We ignore branches with multiple nucleotide mutations to the same codon (this is very rare) because as detailed below our method is only designed to make estimates for mutations that represent single-nucleotide changes from the clade founder.
Note also that the mutation-annotated tree does not include insertion or deletion mutations, and so we only consider (and make estimates for) point mutations.
